Stack Safety is a Security Property

### 2021-04-11 tex/main.tex

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, with consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}
\fi
%
The implementation and configuration of these  stack-safety mechanisms can be complex,
and often involves compromises trading protection for efficiency.
Thus it would be useful to have a precise formal specification for stack
safety, both as a basis for comparing the security guarantees claimed by different enforcement
techniques and for verifying that these claims are met by particular implementations.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
In essence, memory locations within each stack frame are tagged with the identity of
the function activation that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Many alternatives can
be implemented on top of this basic idea, by choosing how to identify activations
and how to maintain the connection between an activation identity and the memory it owns.
These alternatives vary in their
efficiency, but also in the spatial and temporal precision of their security guarantees.
For example, the \emph{Depth Isolation} policy requires a strict mapping from a function
to its frame from the start of the call to its end, and permits the function to write
into that frame. This seems simple and intuitively correct, but it can be expensive to
implement that strict mapping.
On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by
assigning locations to the active function when they are written, which permits
illicit writes outside the function's frame so long as the proper owner of the written
locations does not subsequently read them. The intuition justifying laziness is that any
genuinely dangerous writes will ultimately be detected before they can affect program
behavior.  But we clearly need a more formal foundation for
expressing, comparing and verifying these micro-policy security claims, as well as those of
other enforcement technologies.

@@

In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}.
While our property definitions were developed with micro-policy enforcement in mind, they
agnostic about many implementation details and should be applicable to other
systems with minor adjustment. We define
trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as mutually distrustful
and view the caller as needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
This differs from common instances in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{Thinking about how to express domains, or whether to. I'm not sure it's
  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
{\em Integrity} says that a caller's local data is protected from modification by its callee.
This is vital for local reasoning about the behavior of the caller.
{\em Confidentiality} states that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack.
Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
notion that data outside of live frames should be inaccessible (which LSE does not address).
We also give a simple trace-based formalization of WBCF, which is essential for keeping the
notions of ``caller'' and ``callee'' sensible.

>>>>

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are generally described in terms
of a set of concrete examples of attacks that they can prevent, leaving a
more general characterization to the reader's intuition.
Moreover, the details of these mechanisms can be quite intricate
and often involve trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

@@

\bcp{If we end up adopting my recent proposal for defining subtraces and
  separating WBCF, I think we can be a bit stronger here: We define stack
  safety in IFC terminology based on the same intuitions as their LSE, but
  we consider (and define) WBCF to be a separate thing---desirable, but
  distinct from safety.  (If we make this argument, we'll need to show an
  example of a program that is stack safe but doesn't have WBCF.)}In this
work, we take the (informal) definitions of WBCF and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as
mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
This differs from common instances in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{Thinking about how to express domains, or whether to. I'm not sure it's
  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
{\em Integrity} says that a caller's local data is protected from modification by its callee.
\bcp{What about the ``stepwise'' part?}
This is vital for local reasoning about the behavior of the caller.
{\em Confidentiality} states that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack.
Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
notion that data outside of live frames should be inaccessible (which LSE does not address).
We also give a simple trace-based formalization of WBCF, which is essential for keeping the
notions of ``caller'' and ``callee'' sensible.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Intuitively, memory cells within each stack frame are tagged with the
identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Roessler and Dehon
explore alternative versions of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed, to enable different ways
of maintaining the connection between an activation identity and the memory it owns.
These vary in their efficiency and in the spatial and temporal precision of their security
guarantees.
For example, their \emph{Depth Isolation} policy  requires a strict mapping from a function
to its frame from the start of the call to its end, and permits the function to write
only into that frame.
This seems simple and intuitively correct, but it can be expensive to implement the strict
mapping. Conversely, their \emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. The justification
for this form of laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.
